A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-28 latex/paper.tex
fixed typo
Worker on infracstructure sectcion

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's API for tabular data, is
an overlay of RDDs with similar performance features \VHS{It's probably fine, but the SparkSQL Engine available for Scala and Java do bring an added peformance value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should be mentioned after the API languages such that you only list the APIs relevant to Python (RDD and Dataframes)} Spark's core is written in
Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
the popularity of Python in neuroscience, although the serialization cost of Python
objects to Scala might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches jobs to them\VHS{maybe say tasks instead of jobs, otherwise it sounds like 1 job/worker}. A job is divided
into stages to be processed onto the workers. Each stage's operation is represented
as a high-level task in the computation graph. The Spark standalone scheduler uses a
FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
increasing the page hit rate. As our applications only have one job, the LIFO
(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark and Dask opt to complete a chain of dependencies rather than start a bunch of ``new chains''. And in this context, LIFO is great for the page hit rate, so you might want to reword the prior sentence as well}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as cluster mode in not available in PySpark. We used Apache Spark v2.4.0.

@@

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a programming
abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
parallel composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.
\VHS{maybe mention which one use threads and which use processes}

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8
 threads. We used Dask's default configuration that uses all the available
 memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
 configured Spark to use 1 executor per worker and 25~GB of memory per
 executor, to leave 5~GB for off-heap. \TG{Check spill to disk}

 One cloud instance did not host any worker, and had a 2~TB disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. We configured the Spark driver to use 25~GB of memory, and
 used the default configuration for the master.

@@

We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
brain with voxel intensities ranging from 0 to 65,535. The original data is
stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
resolution of \SI{40}{\micro\metre}. We converted the blocks into the
\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
a total data size of 76~GiB \TG{Check if GiB or GB}. 
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
2.5GB, 0.6GB, and
0.1GB \TG{check sizes}, using the sam~\cite{sam} library.

We also used the dataset provided by the Consortium for Reliability and
Reproducibility
(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on
\href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is 408.4~GB, containing anatomical, diffusion and functional
images of 1,397 subjects acquired in 29 sites. Among these, we only used
data from \TG{X} subjects, representing \TG{Y}GB overall (\TG{Z}GB by
subject on average).

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage[binary-units=true]{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's API for tabular data,
is an overlay of RDDs with similar performance features \VHS{It's probably fine, but
the SparkSQL Engine available for Scala and Java do bring an added peformance
value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for
DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should
be mentioned after the API languages such that you only list the APIs relevant to
Python (RDD and Dataframes)} Spark's core is written in Scala, but it also has a
Java, an R, and a Python API called PySpark. \MD{Sentence is ambiguous? could think
that all three APIs are named PySpark} We used PySpark, given the popularity of
Python in neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default scheduler, Spark
standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches tasks to them. A job is divided
into stages to be excuted in a different process onto the workers. Each stage's
operation is represented as a high-level task in the computation graph. The Spark
standalone scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
Processing jobs in order helps in increasing the page hit rate. As our applications
only have one job, the LIFO (Last-In-First-Out) policy used for the stages is more
pertinent \MD{Find better phrasing} \TG{refer to the paper (\MD{Not sure which one})}
\VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank.
Basically both Spark and Dask opt to complete a chain of dependencies rather than
start a bunch of ``new chains''. And in this context, LIFO is great for the page hit
rate, so you might want to reword the prior sentence as well}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.

@@

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It use multi-process to enable
parallelism. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. An
\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the processing
of large arrays. It provides a distributed clone of the popular NumPy library. A
\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.
\VHS{maybe mention which one use threads and which use processes}

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and 25~GB of memory per executor, to leave 5~GB
 for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
 default configuration for the master. We used the default confirguration for worker
 memory management: at 60\% it spill data to disk, and 50\% of that amount is
 reserved for storage and is immuned from eviction.

 One cloud instance did not host any worker, and had a 2~TB disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data is evicted to the NFS.

@@

We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
brain with voxel intensities ranging from 0 to 65,535. The original data is
stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
resolution of \SI{40}{\micro\metre}. We converted the blocks into the
\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
a total data size of \SI{81}{\giga\byte}. 
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
\SI{2.7}{\giga\byte}, \SI{0.648}{\giga\byte}, and
\SI{0.108}{\giga\byte}, using the sam~\cite{sam} library.

We also used the dataset provided by the Consortium for Reliability and
Reproducibility
(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on
\href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is 408.4~GB, containing anatomical, diffusion and functional
images of 1,397 subjects acquired in 29 sites. Among these, we only used
data from \TG{X} subjects, representing \TG{Y}GB overall (\TG{Z}GB by
subject on average).
\MD{X, Y, Z values TBD after testing the makespan with different number of subjects.}
